alternativeRSA

### 2016-01-29 models/RSA_alternativeCostFormulation.Rmd
added generic None (all prob mass under one star runs

  scales <- unique(data$scale)
  for (s in scales) {
      if (numAlts == 5) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 mid = mid / sum(mid),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, mid, low1, low2)
      } else if (numAlts == 4) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, low1, low2)
      } else if (numAlts == 2) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi = hi / sum(hi),
                 low = low / sum(low)) %>%
          select(hi, low)
      }
    stars <- subset(data, scale == s)$stars
    degrees <- as.character(subset(data, scale == s)$degree)
    currCosts <- unique(subset(data, scale == s) %>% select(degree, cost))
    priors <- data %>% select(c(stars, prior.p)) %>% head(., n=5)

@@

Data set-up
```{r}
# All scales studied
# ------------------
scales <- c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
# Priors
# ------
unifPriors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
empPriors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
yelpPriors <- data.frame(stars=1:5,
                         prior.p =
                           c(0.090190981, 0.105689431, 0.162783722, 0.304569543, 0.336766323))
# Cost / rewards
# --------------
emptyCosts5 <- data.frame(scale = c(rep("good_excellent", 5),
                                    rep("liked_loved", 5),
                                    rep("memorable_unforgettable", 5),
                                    rep("palatable_delicious", 5),
                                    rep("some_all", 5)),
                          degree = rep(c("hi1", "hi2", "mid", "low1", "low2"), 5),
                          newCount = rep(0.0, 25),
                          stringsAsFactors = FALSE)
emptyCosts4 <- data.frame(scale = c(rep("good_excellent", 4),
                                    rep("liked_loved", 4),
                                    rep("memorable_unforgettable", 4),
                                    rep("palatable_delicious", 4),
                                    rep("some_all", 4)),
                          degree = rep(c("hi1", "hi2", "low1", "low2"), 5),
                          newCount = rep(0.0, 20),
                          stringsAsFactors = FALSE)
emptyCosts2 <- data.frame(scale = c(rep("good_excellent", 2),
                                    rep("liked_loved", 2),
                                    rep("memorable_unforgettable", 2),
                                    rep("palatable_delicious", 2),
                                    rep("some_all", 2)),
                          degree = rep(c("hi", "low"), 5),
                          newCount = rep(0.0, 10),
                          stringsAsFactors = FALSE)

@@

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8Compliance.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)

@@

  for (s in scales) {
    saliences[saliences$scale == s, ]$cost <-
      (-1 * saliences[saliences$scale == s, ]$newCount) / NUM_ALTERNATIVES_E9
    saliences[saliences$scale == s, ]$cost <-
      -1 * normedVec(saliences[saliences$scale == s, ]$cost)
  }
  saliences[, c("scale", "degree", "cost")]
}
############# NEW SALIENCES 1/20/16 #########

costs <- normNegSalienceCosts(salienceCosts5)
priors <- unifPriors
data <- left_join(speakerL12, costs) %>%
  left_join(., priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, prior.p, cost) %>%
  group_by(scale)

@@

  for (s in unique(predDf$scale)) {
    if (study == "e12") {
      if (s == "some_all") {
        tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "mid")) %>%
          mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else if (study == "e10") {
        if (s == "some_all") {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "low1")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else { ## study e6
      tempDf <- subset(predDf, scale == s)
    }

@@

# Alpha Tuning - maximixize correlation
# -------------------------------------
tuneAlphas <- function(modelData, humanDf, alphas = seq(1, 12), numAlts = 5, study = "e12"){
  sapply(alphas, function(i) {
    predDf <- runFull(modelData, calcFn = 2, alpha = i, numAlts = numAlts)
    mergedData <- matchModelHumanData(predDf, listener, study = study)
    cor(mergedData$preds, mergedData$listener.p)
  })
}
# tuneAlphas(modelData = data, humanDf = listener, numAlts = 4)

@@

Symmetric scale analysis
```{r}
# Set-up data

completeModelRun <- function(speaker, listener, priors, costs, numAlts) {
  results <- list() ## Will return this
  newCosts <- normNegSalienceCosts(costs)

@@

  studyNum <- ifelse(numAlts == 5, "e12", ifelse(numAlts == 4, "e10", "e6"))
  # Tune alpha
  bestAlpha <- which.max(tuneAlphas(data, listener, numAlts = numAlts, study = studyNum))
  # Run model
  modelData <- runFull(data, calcFn = 2, alpha = bestAlpha, numAlts = numAlts)
  # Final data set for plotting / model fit

@@

Run complete analyses and output to csv
```{r}
# Run analyses
symmetricScalerun <- completeModelRun(speakerL12, listener, unifPriors, emptyCosts5, numAlts = 5)
symmetricScalerun[[4]]
negAltsRun <- completeModelRun(speakerL10, listener, unifPriors, emptyCosts4, numAlts = 4)
negAltsRun[[4]]
entailmentRun <- completeModelRun(speakerL08, listener, unifPriors, emptyCosts2, numAlts = 2)
entailmentRun[[4]]

listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)
symmetricData <- symmetricScalerun[[1]]
symmetricData$numAlts = 5
negAltsData <- negAltsRun[[1]]
negAltsData$numAlts = 4
entailmentData <- entailmentRun[[1]]
entailmentData$numAlts = 2

@@

# combine with e11 prag listener nums
aggregateData <- rbind(symmetricData, negAltsData, entailmentData)
listener2 <- listener2[(listener2$scale == "some_all" & (listener2$degree == "low1" | listener2$degree == "hi1")) |
                         (listener2$scale != "some_all" & (listener2$degree == "hi1" | listener2$degree == "hi2")), ]
listener2$degree <- as.character(ifelse(listener2$degree == "hi1", "hi", "low"))
names(listener2) <- c("degree", "scale", "stars", "listener.p2", "total")
aggregateData <- left_join(aggregateData, listener2[, -5], by=c("degree", "stars", "scale")) %>%
  mutate(listener.p2 = ifelse(is.na(listener.p2), 0, listener.p2))

# record correlations for both studies
modelRunData <- aggregateData
symmetricE6cor <- cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p)
symmetricE11cor <-cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p2)
negAltsE6cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p)
negAltsE11cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p2)
entailmentE6cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p)
entailmentE11cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p2)
modelRunData$corE6 <- ifelse(modelRunData$numAlts == 5, symmetricE6cor,
                             ifelse(modelRunData$numAlts == 4, negAltsE6cor, entailmentE6cor))
modelRunData$corE11 <- ifelse(modelRunData$numAlts == 5, symmetricE11cor,
                             ifelse(modelRunData$numAlts == 4, negAltsE11cor, entailmentE11cor))
modelRunData$alphas <- ifelse(modelRunData$numAlts == 5, symmetricScalerun[[3]],
                             ifelse(modelRunData$numAlts == 4, negAltsRun[[3]], entailmentRun[[3]]))

>>>>

  scales <- unique(data$scale)
  for (s in scales) {
      if (numAlts == 5) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 mid = mid / sum(mid),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, mid, low1, low2)
      } else if (numAlts == 4) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, low1, low2)
      } else if (numAlts == 3) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi = hi / sum(hi),
                 low = low / sum(low),
                 none = none / sum(none)) %>%
          select(hi, low, none)
      } else if (numAlts == 2) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi = hi / sum(hi),
                 low = low / sum(low)) %>%
          select(hi, low)
      }
    stars <- subset(data, scale == s)$stars
    degrees <- as.character(subset(data, scale == s)$degree)
    currCosts <- unique(subset(data, scale == s) %>% select(degree, cost))
    priors <- data %>% select(c(stars, prior.p)) %>% head(., n=5)

@@

Data set-up
```{r}
# All scales studied
# ------------------
scales <- c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
# Priors
# ------
unifPriors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
empPriors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
yelpPriors <- data.frame(stars=1:5,
                         prior.p =
                           c(0.090190981, 0.105689431, 0.162783722, 0.304569543, 0.336766323))
# Cost / rewards
# --------------
# for five alts
emptyCosts5 <- data.frame(scale = c(rep("good_excellent", 5),
                                    rep("liked_loved", 5),
                                    rep("memorable_unforgettable", 5),
                                    rep("palatable_delicious", 5),
                                    rep("some_all", 5)),
                          degree = rep(c("hi1", "hi2", "mid", "low1", "low2"), 5),
                          newCount = rep(0.0, 25),
                          stringsAsFactors = FALSE)
# for four alts
emptyCosts4 <- data.frame(scale = c(rep("good_excellent", 4),
                                    rep("liked_loved", 4),
                                    rep("memorable_unforgettable", 4),
                                    rep("palatable_delicious", 4),
                                    rep("some_all", 4)),
                          degree = rep(c("hi1", "hi2", "low1", "low2"), 5),
                          newCount = rep(0.0, 20),
                          stringsAsFactors = FALSE)
# Used with generic none run
emptyCosts3 <- data.frame(scale = c(rep("good_excellent", 3),
                                    rep("liked_loved", 3),
                                    rep("memorable_unforgettable", 3),
                                    rep("palatable_delicious", 3),
                                    rep("some_all", 3)),
                          degree = rep(c("hi", "low", "none"), 5),
                          newCount = rep(0.0, 15),
                          stringsAsFactors = FALSE)
# for two-alts
emptyCosts2 <- data.frame(scale = c(rep("good_excellent", 2),
                                    rep("liked_loved", 2),
                                    rep("memorable_unforgettable", 2),
                                    rep("palatable_delicious", 2),
                                    rep("some_all", 2)),
                          degree = rep(c("hi", "low"), 5),
                          newCount = rep(0.0, 10),
                          stringsAsFactors = FALSE)

@@

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerGenericNone <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8_genericNone.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8Compliance.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)

@@

  for (s in scales) {
    saliences[saliences$scale == s, ]$cost <-
      saliences[saliences$scale == s, ]$newCount / NUM_ALTERNATIVES_E9

    saliences[saliences$scale == s, ]$cost <- 1 / log(saliences[saliences$scale == s, ]$cost)
  }
  saliences[, c("scale", "degree", "cost")]
}
############# NEW SALIENCES 1/20/16 #########

# costs <- normNegSalienceCosts(salienceCosts5)
# priors <- unifPriors
# data <- left_join(speakerL12, costs) %>%
#   left_join(., priors) %>%
#   rowwise %>%
#   select(scale, degree, stars, speaker.p, prior.p, cost) %>%
#   group_by(scale)

@@

  for (s in unique(predDf$scale)) {
    if (study == "e12") {
      if (s == "some_all") {
        tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "mid")) %>%
          mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else if (study == "e10") {
        if (s == "some_all") {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "low1")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else if (study == "genericNone") {
      tempDf <- subset(predDf, scale == s & degree != "none")
    } else { ## study e6
      tempDf <- subset(predDf, scale == s)
    }

@@

# Alpha Tuning - maximixize correlation
# -------------------------------------
tuneAlphas <- function(modelData, humanDf, alphas = seq(1, 12), numAlts = 5, study = "e12"){
  sapply(alphas, function(i) {
    predDf <- runFull(modelData, calcFn = 2, alpha = i, numAlts = numAlts)
    # Watch for generic none run
    if (study == "genericNone") {
      mergedData <- matchModelHumanData(subset(predDf, degree != "none"), listener, study = study)
    } else {
      mergedData <- matchModelHumanData(predDf, listener, study = study)
    }
    cor(mergedData$preds, mergedData$listener.p)
  })
}
# tuneAlphas(modelData = data, humanDf = listener, numAlts = 4)

@@

Run full model analysis
```{r}
# completeModelRun()
# ------------------
completeModelRun <- function(speaker, listener, priors, costs, numAlts) {
  results <- list() ## Will return this
  newCosts <- normNegSalienceCosts(costs)

@@

  studyNum <- ifelse(numAlts == 5, "e12",
                     ifelse(numAlts == 4, "e10",
                            ifelse(numAlts == 3, "genericNone", "e6")))
  # Tune alpha
  bestAlpha <- which.max(tuneAlphas(data, listener, numAlts = numAlts, study = studyNum))
  # Run model
  modelData <- runFull(data, calcFn = 2, alpha = bestAlpha, numAlts = numAlts)
  # Final data set for plotting / model fit

@@

Run complete analyses and output to csv
```{r}
# Run analyses
fiveAlts <- completeModelRun(speakerL12, listener, unifPriors, emptyCosts5, numAlts = 5)
fiveAlts[[4]]
fourAlts <- completeModelRun(speakerL10, listener, unifPriors, emptyCosts4, numAlts = 4)
fourAlts[[4]]
twoAltsGenericNone <- completeModelRun(speakerGenericNone, listener, unifPriors, emptyCosts3, numAlts = 3)
twoAltsGenericNone[[4]]
twoAlts <- completeModelRun(speakerL08, listener, unifPriors, emptyCosts2, numAlts = 2)
twoAlts[[4]]

listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)
fiveAltsData <- fiveAlts[[1]]
fiveAltsData$numAlts = 5
fourAltsData <- fourAlts[[1]]
fourAltsData$numAlts = 4
twoAltsGenericNoneData <- twoAltsGenericNone[[1]]
twoAltsGenericNoneData$numAlts = 3
twoAltsData <- twoAlts[[1]]
twoAltsData$numAlts = 2

@@

# combine with e11 prag listener nums
aggregateData <- rbind(fiveAltsData, fourAltsData, twoAltsGenericNoneData, twoAltsData)
listener2 <- listener2[(listener2$scale == "some_all" & (listener2$degree == "low1" | listener2$degree == "hi1")) |
                         (listener2$scale != "some_all" & (listener2$degree == "hi1" | listener2$degree == "hi2")), ]
listener2$degree <- as.character(ifelse(listener2$degree == "hi1", "hi", "low"))
names(listener2) <- c("degree", "scale", "stars", "listener.p2", "total")
aggregateData <- left_join(aggregateData, listener2[, -5], by=c("degree", "stars", "scale")) %>%
  mutate(listener.p2 = ifelse(is.na(listener.p2), 0, listener.p2))

# record correlations for both studies
modelRunData <- aggregateData
fiveAltsE6cor <- cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p)
fiveAltsE11cor <-cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p2)
fourAltsE6cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p)
fourAltsE11cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p2)
twoAltsGenericNoneE6cor <-
  cor(subset(modelRunData, numAlts == 3)$preds, subset(modelRunData, numAlts == 3)$listener.p)
twoAltsGenericNoneE11cor <-
  cor(subset(modelRunData, numAlts == 3)$preds, subset(modelRunData, numAlts == 3)$listener.p2)
twoAltsE6cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p)
twoAltsE11cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p2)

modelRunData$corE6 <- ifelse(modelRunData$numAlts == 5, fiveAltsE6cor,
                             ifelse(modelRunData$numAlts == 4, fourAltsE6cor,
                                    ifelse(modelRunData$numAlts == 3, twoAltsGenericNoneE6cor, twoAltsE6cor)))
modelRunData$corE11 <- ifelse(modelRunData$numAlts == 5, fiveAltsE11cor,
                             ifelse(modelRunData$numAlts == 4, fourAltsE11cor,
                                    ifelse(modelRunData$numAlts == 3, twoAltsGenericNoneE11cor, twoAltsE11cor)))
modelRunData$alphas <- ifelse(modelRunData$numAlts == 5, fiveAlts[[3]],
                             ifelse(modelRunData$numAlts == 4, fourAlts[[3]],
                                    ifelse(modelRunData$numAlts == 3, twoAltsGenericNone[[3]], twoAlts[[3]])))
